### 2010-12-20 Documentation/atropos/atropos.tex
ENH: Update to introduction, theory and experiment sections.  Adding Hammers figure.  BWeb results next.
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@707 fd2da0a4-d968-48df-9cb5-d8c496287cbf

%\documentclass[11pt,english]{article}
%
%% Set page margins correctly
%\usepackage{geometry}
%%\geometry{letterpaper,top=0.5in,left=0.5in,bottom=0.5in,top=0.5in,headsep=6pt,footskip=18pt}
%\geometry{letterpaper,top=0.5in,left=0.5in,bottom=0.5in,top=0.5in,headsep=6pt,footskip=18pt}
%
%% Use fancy header style
%\usepackage{fancyhdr}
%\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.75pt}
%\renewcommand{\footrulewidth}{0.75pt}
%\lhead{\footnotesize Principal Investigator/Program Director (Last, First, Middle): }
%\rhead{Yushkevich, Paul A.} \lfoot{\footnotesize PHS 398/2590 (Rev. 04/06)} \rfoot{\footnotesize
%\textbf Continuation Format Page} \cfoot{\footnotesize Page \underline{~~\thepage~~}}
%
%% Use pslatex fonts
%\usepackage{pslatex}
%\renewcommand{\familydefault}{\sfdefault}
%\renewcommand{\baselinestretch}{.9}
%

% START JUNK
\documentclass[11pt,english]{article}

% Set page margins correctly
\usepackage{geometry}
\usepackage{url}
\geometry{letterpaper,top=1.0in,left=1.0in,bottom=1.0in,top=1.0in,headsep=6pt,footskip=18pt}

\usepackage{lscape}
\usepackage[round,semicolon,authoryear,]{natbib}

% Use fancy header style
\usepackage{fancyhdr}
\pagestyle{empty}
\renewcommand{\headrulewidth}{0.75pt}
\renewcommand{\footrulewidth}{0.75pt}
%\lhead{\footnotesize Principal Investigator/Program Director (Last, First, Middle): }
%\rhead{Gee, James C.} \lfoot{\footnotesize PHS398 (06/07)} \rfoot{\footnotesize
%\textbf Continuation Format Page} \cfoot{\footnotesize Page \underline{~~\thepage~~}}

\usepackage{setspace}
\usepackage{listings}
\usepackage{float}

\floatstyle{plain}
\newfloat{command}{thp}{lop}
\floatname{command}{Command}

% \doublespacing

% Use pslatex fonts
%\usepackage[T1]{fontenc}
%\usepackage{mathptmx}

% END JUNK

% All other packages
\usepackage{boxedminipage}
\usepackage{graphicx}
\usepackage{amsmath,amsfonts}
\usepackage{babel,verbatim}
\usepackage{enumerate}
%\usepackage{dsfont}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage[compact]{titlesec}
\usepackage[usenames]{color}
\usepackage{ulem}
\usepackage{multirow,booktabs,ctable,array}

%GATHER{../../../shared/bibtex/biblio.bib}
%GATHER{../../../shared/bibtex/brainmri.bib}
%GATHER{../../../shared/bibtex/registration.bib}
\graphicspath{{./Figures/}
                          }

%\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}

\long\def\symbolfootnote[#1]#2{\begingroup%
\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

    \usepackage{color}

    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
    \definecolor{listnumbers}{gray}{0.65}
    \definecolor{listlightgray}{gray}{0.955}
    \definecolor{listwhite}{gray}{1.0}

\newcommand{\lstsetcpp}
{
\lstset{frame = tb,
        framerule = 0.25pt,
        float,
        fontadjust,
        backgroundcolor={\color{listlightgray}},
        basicstyle = {\ttfamily\scriptsize},
        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
        identifierstyle = {\ttfamily},
        commentstyle = {\ttfamily\color{listcomment}\textit},
        stringstyle = {\ttfamily},
        showstringspaces = false,
        showtabs = false,
        numbers = none,
        numbersep = 6pt,
       numberstyle={\ttfamily\color{listnumbers}},
        tabsize = 2,
        language=[ANSI]C++,
        floatplacement=!h,
        caption={\small \baselineskip 12pt Atropos short command line menu which is invoked using the `{\ttfamily -h}' option. 
        The expanded menu, which provides additional details regarding the possible parameters and usage 
        options, is elicited using the `{\ttfamily {-}{-}help}' option.},
        captionpos=b,
        label=listing:command
        }
}

\newcommand{\lstsetcppnfour}
{
\lstset{frame = tb,
        framerule = 0.25pt,
        float,
        fontadjust,
        backgroundcolor={\color{listlightgray}},
        basicstyle = {\ttfamily\scriptsize},
        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
        identifierstyle = {\ttfamily},
        commentstyle = {\ttfamily\color{listcomment}\textit},
        stringstyle = {\ttfamily},
        showstringspaces = false,
        showtabs = false,
        numbers = none,
        numbersep = 6pt,
        numberstyle={\ttfamily\color{listnumbers}},
        tabsize = 2,
        language=[ANSI]C++,
        floatplacement=!h,
        caption={\small \baselineskip 12pt N4 short command line menu which is invoked using the `{\ttfamily -h}' option. 
        The expanded menu, which provides additional details regarding the possible parameters and usage 
        options, is elicited using the `{\ttfamily {-}{-}help}' option.  },
        captionpos=b,
        label=listing:n4
        }
}

%\renewcommand{\topfraction}{0.85}
%\renewcommand{\textfraction}{0.1}
%\renewcommand{\floatpagefraction}{0.75}

% Different font in captions
%\newcommand{\captionfonts}{\small}
%\makeatletter  % Allow the use of @ in command names
%\long\def\@makecaption#1#2{%
%  \vskip\abovecaptionskip
%  \sbox\@tempboxa{{\captionfonts #1: #2}}%
%  \ifdim \wd\@tempboxa >\hsize
%    {\captionfonts #1: #2\par}
%  \else
%    \hbox to\hsize{\hfil\box\@tempboxa\hfil}%
%  \fi
%  \vskip\belowcaptionskip}
%\makeatother   % Cancel the effect of \makeatletter

%Automated Segmentation of Ventilation Defects on $^3$He Lung MRI}
%\date{}
%\author{ Nicholas J. Tustison, DSc,$^{1*}$ Talissa A. Altes, MD,$^2$ Eduard E. de Lange, MD,$^2$
%  Brian B. Avants, PhD, $^1$ John P. Mugler III, PhD,$^2$ and James C. Gee, PhD,$^1$ \\
%  $^1$ Penn Image Computing and Science Laboratory, University of Pennsylvania, Philadelphia, Pennsylvania,  USA.\\
%  $^2$ Department of Radiology, University of Virginia, Charlottesville, Virginia, USA}
%\address{3600 Market Street, Suite 370 \\
%                 Philadelphia, PA  19104\\
%                 tustison@picsl.upenn.edu}
%\advisor{James Gee}
\begin{document}

\normalem

\vspace*{5cm}

\begin{center}
{\Large \bf An Open Source Framework for Multivariate $n$-Tissue
  Segmentation and Evaluation Using Open Data Sets} \\
\vspace*{0.5cm}
{\normalsize Brian B.~Avants$^{1*}$,  Nicholas J.~Tustison$^2$%
\symbolfootnote[1]{
  The first two authors contributed equally to this work.
}, 
Philip A.~Cook$^1$, Jue Woo$^1$, and James C.~Gee$^1$} \\
\begin{singlespace} 
{\scriptsize  $^1$ Penn Image Computing and Science Laboratory, University of Pennsylvania, Philadelphia, Pennsylvania,  USA.\\
  $^2$ Department of Radiology, University of Virginia, Charlottesville, Virginia, USA}
\end{singlespace}
\end{center}

\vfill

\begin{singlespace} 
\scriptsize
\flushleft
%\line(1, 0){250} \\
{\bf Atropos:  $n$-Tissue Segmentation}\\
Corresponding author: \\
Brian B. Avants\\
3600 Market Street, Suite 370\\
Philadelphia, PA  19104\\
avants@picsl.upenn.edu\\
\end{singlespace}

%\clearpage
%
%\vspace*{7cm}
%
%\begin{center}
%{\Large \bf An Open Source Framework for Multivariate $n$-Tissue Segmentation and Volumetric Cortical Parcellation:  Evaluation Using the BrainWeb, NIREP, and LPBA Data Sets}\\
%\end{center}

\clearpage

\begin{abstract} 
%Neuroanatomical coordinate systems are essential for the
%interpretation of structural and functional imaging studies.
%However, manual delineation of the sulco-gyral complex is time
%consuming and prone to variability given cortical complexity.  This
%work describes an open source, image-based algorithmic approach to
%cortical parcellation which uses training data to propagate
%structural labelings to individual images.  Along with other
%neuroinformatics tools publicly distributed in our ANTs (Advanced Normalization Tools)%
We introduce {\em Atropos}, an ITK-based multivariate $n$-class
Expectation Maximization (EM) segmentation algorithm based on either parametric or non-parametric finite mixture modeling (FMM) of the
class intensities.  Atropos, available in ANTs
\footnote{
http://www.picsl.upenn.edu/ANTs
}, is capable of incorporating spatial prior probability maps,
(sparse) prior label maps and/or Markov Random Field (MRF) modeling.  Atropos has also been efficiently implemented to handle
large quantities of possible labelings (in the experimental section,
we use up to 68 classes) with a minimal memory footprint.  This work
describes the technical and implementation aspects of Atropos and
evaluates its performance on two different ground-truth datasets.
First, we use the BrainWeb dataset from Montreal Neurological
Institute to evaluate three tissue segmentation performance via (1)
K-means segmentation without use of template data; (2) MRF segmentation with initialization by prior probability maps
derived from a group template; (3) Prior-based segmentation with use
of spatial prior probability maps derived from a group template.  We
also evaluate Atropos performance by using spatial priors to drive a 68 class EM segmentation
problem defined by the Hammers atlas from University College London.  These evaluation studies, combined with
illustrative examples that excercise Atropos options, demonstrate both
performance and wide applicability of this new platform-independent 
open source segmentation tool.
%Component (3) corresponds to the most
%realistic clinical case, while component (2) eliminates the confound
%of defining the cortex itself and focuses only on the parcellation
%problem.  Consistent with the open source dissemination of ANTs, we
%make publicly available all scripts and code from the evaluation.
\end{abstract}

%\begin{keyword}
%segmentation \sep expectation maximization \sep spatial prior \sep brain 
%\end{keyword}

\clearpage

\section{Introduction} As medical image acquisition technology has advanced,
concomitant investment has been made towards adapting classification
techniques for neuroanatomy.  Some of the earliest work appropriated NASA
satellite image processing software for statistical classification of
head tissues in 2-D MR images \citep{Vannier1985}.  A proliferation of
techniques ensued with increasing sophistication in both core
methodology and degree of refinement for specific problems.  The
chronology of progress in segmentation may be tracked through both
technical reviews
\citep{Bezdek1993,Pal1993,Clarke1995,Pham2000,Viergever2001,Suri2002,Duncan2004,Balafar2010}
and evaluation studies
\citep[e.g.][]{Cuadra2005,Zaidi2006,Klauschen2009,Boer2010}.

The problem of accurately delineating the white matter, grey matter
and cerebrospinal fluid of the primate brain continuously spurs
technical development in segmentation.  Following \cite{Vannier1985},
many researchers adopted statistical methods for $n$-tissue anatomical
brain segmentation.  Given the ``missing data'' aspect of this
problem, a natural optimization strategy is the use of the
Expectation-Maximization (EM) framework \citep{Dempster1977}.  The
work described in \cite{Wells1996} was one of the first to use EM for
finding a locally optimal solution by iterating between bias field
estimation and tissue segmentation.  A core component of this work was
explicit modeling of the tissue intensity values as normal
distributions \citep{Cline1990} for both 2-D univariate simulated data
and T1 coronal images, which continues to find utility in contemporary
developments.  A secondary component was an extended nonparametric
approach, also influenced by earlier work \citep{Kikinis1992}, where
Parzen windowing is used to model the tissue intensity distribution
omitting consideration of the underlying bias field.  Although
technically not an EM-based algorithm, the robustness of the latter
approach has motivated its continued use even more recently \citep[e.g.][]{Weisenfeld2009}.

Subsequent development included the use of Markov Random Field (MRF)  modeling \citep{Geman1984}
to regularize the classification results \citep{Held1997} with later work adding heuristics concerning neuroanatomy to prevent 
over-regularization and the resulting loss of fine structural details \citep{Leemput1999,Leemput1999a}.  
A more formalized integration of generic MRF spatial priors was employed in the work of \cite{Zhang2001}, 
commonly referred to as FAST (FMRIB's Automated Segmentation Tool), which is in widespread use
given its public availability and good performance.  More recently, a uniform distribution of local MRFs within the brain volume and their subsequent integration into a global solution has been proposed obviating the need for an 
explicit bias correction solution \citep{Scherrer2009}.

EM algorithms are sensitive to local optima. Consequently, a good initialization
strategy is critical.  Common low-level initialization strategies include uniform
probability assignment \citep{Wells1996}, Otsu thresholding
\citep{Zhang2001}, and K-means clustering \citep{Pappas1992}.  More
sophisticated low-level initialization schemes include that of
\cite{Greenspan2006} in which a dense spatial distribution of
Gaussians is used to capture the complex neuroanatomical layout with
subsequent processing used to conjoin subsets of Gaussians belonging
to the same tissue classes.  Recently, anatomically specific
strategies for incorporating prior knowledge into the segmentation
solution have included the use of spatial prior probability maps of
the various structures of interest
\citep{Leemput1999a,Marroquin2002,Ashburner2005}.  These spatial prior
probability maps can also be used to provide an initial segmentation.
Related technological developments model partial
volume effects for increased accuracy in brain segmentation
\citep{Ruan2000,Ballester2002,Leemput2003}.

A general trend towards more integrative neuroanatomical image processing led to the work described in
\cite{Ashburner2005} which is publicly available within SPM5, a
large-scale Matlab module in which registration, segmentation, and
bias field correction can be simultaneously modeled within a single
optimization scheme. The roots of this very popular software package
stem back to early work by Karl Friston in which the basis for
statistical parametric mapping was developed \citep{Friston1990} and
subsequently provided to the research community. A similar holistic
approach to brain processing was provided in \cite{Pohl2006} in which
segmentation and registration parameters were optimized simultaneously
while casting the inhomogeneity model parameters of \cite{Wells1996}
as nuisance variables.  Continued work involved recursive parcellation
of the brain volume by considering sub-structures in a hierarchical
manner \citep{Pohl2007}.  An implementation is provided in 3D slicer
\citep{Pieper2006}---an open source package with developmental
contributions from multiple agencies including both private and
academic institutions. Researchers in aging often focus on
accurately segmenting the T1 MRI of elderly controls and subjects
suffering from neurodegeneration, for instance, via the SIENA algorithm
\citep{Smith2007}.  A recent evaluation study 
compared kNN segmentation, SPM Unified Segmentation and SIENA and
found different performance
characteristics under different evaluation criteria
\citep{Bresser2011}.  \cite{Klauschen2009} had similar findings when
comparing SPM5, FSL and FreeSurfer.  These studies suggest that no
single method performs best under every measurement and, along with the No
Free Lunch theorem, highlight the need for segmentation tools that are tunable for different
problems and research goals.

Related neuroanatomical research concerns the selection of geometric
features of the cortex \citep[e.g.][]{Goualher1999} with aims at
understanding the functional-anatomical relationship of the human
brain. Recent endeavors produce a dense cortical labeling in which
every point of the cortex is classified, i.e. a cortical parcellation
\citep{Fischl2004,Heckemann2006,Destrieux2010}.  Given the extreme
time-consumption and tedium associated with manual effort towards the
latter, various algorithms have been proposed of which a small subset
has been publicly availed to the research community such as the
popular software package known as Freesurfer
\citep{Dale1999,Fischl1999,Fischl2004}.  In contrast to the volumetric
approach detailed in this work, Freesurfer is primarily a
surface-based technique in which the brain structures such as the
grey-white matter interface and pial surfaces are processed, analyzed,
and displayed as tessellated surfaces  \cite{Dale1999,Fischl1999}.
Advantages of surface-based approaches include the ability to map
processed neuroanatomy to simple geometric primitives such as spheres
or planes and the ease of including topological constraints in the
analysis workflow.  These types of methods, as well as Klein's
Mindboggle \cite{Klein2005}, would usually follow an initial segmentation by a volumetric method such as Atropos.

%{\color{red}{{\em Needs work}:  Mindboggle, Heckemann  One advantage
%of the pipeline we propose is that we parcellate cortex entirely in
%the image space, thus avoiding the difficulty of transferring labels
%from the mesh space back to the image space---a problem that is a confound of surface-based methods\citep{Klein2010}.}}

Our open source segmentation tool, which we have dubbed {\em
Atropos}, 
\footnote{ Atropos is one of the three Fates from Greek
mythology characterized by her dreaded shears used to decide the
destiny of each mortal.  Also, consistent with the entomological motif
of our ANTs toolkit, {\it Acherontia atropos} is a species of large
moth known for the skull-like pattern visible on its thorax.  }
efficiently and flexibly implements a $n$-tissue paradigm for
voxel-based image segmentation.  Atropos allows users to harness its
generalized EM algorithm for standard tissue classification of
the brain into gray matter, white matter and cerebrospinal fluid.
Atropos equally allows incarnations that use EM to simultaneously
maximize the posterior probabilities of many classes, for instance,
when parcellating the brain into hemispheres, cortical regions and
deep brain structures such as amygdala, hippocampus and
thalamus.  Atropos contains features of its predecessors for
performing $n$-tissue segmentation including imposition of prior
information in the form of MRFs and template-based spatial prior
probability maps as well as weighted combinations of these terms.  We
also borrow an idea from \cite{Boykov2004} and use sparse spatial
priors to provide initialization and boundary conditions for Atropos
EM segmentation in a semi-interactive manner.  In short, Atropos seeks
to provide a toolbox of segmentation approaches that may be modified,
tuned and refined by the user for different problems.

We also allow Atropos to address the bias field issue by incorporating
the recently developed N4 bias correction algorithm
\citep{Tustison2010} in an adaptive manner.  Coupled with the
registration \citep{Avants2010b} and template building strategies
\citep{Avants2010} already included in the ANTs toolkit, Atropos is a
versatile and powerful software tool which touches multiple aspects of
our brain processing pipeline from brain extraction
\citep{Avants2010a} to grey matter/white matter/cerebrospinal fluid
segmentation to label fusion and propagation for achieving a refined
cortical parcellation based on both prior information and
subject-specific image data.

The Atropos algorithm is multi-platform, free, stand-alone and able to
manage a wide range of scalar and vector image datatypes.  To highlight the
value of this open-source contribution, we performed a search of
software attributes on NITRC and found that as of November 2010 no
stand-alone EM methods are currently listed.
Atropos's memory efficiency enables EM-segmentation of many classes at
once, a capability that is due to a sparse internal representation of
prior probability images that minimizes the need for random access
memory.  We evaluate EM-based classification of $1 mm^3$ T1 MR
neuroimages into 68 neuronatomical regions to illustrate the practical
value of the low-memory implementation within this paper.  Atropos
also allows multivariate image features to drive the data term, relevant when more than one view of anatomy aids
segmentation, as in neonatal brain tissue classification
\cite{Weisenfeld2009}.  Although Atropos may be applied to
multivariate data from arbitrary modalities, we limit our evaluation
to tissue-classification in T1 neuroimaging in part due to the
abundance of ``gold-standard'' data for this modality.  Consistent
with our advocacy of open science (not to mention the facilitation of
analysis due to accessibility) we also only use publicly available
data sets.  For this reason, all results in this paper are
reproducible with the caveat that users may require some guidance from
the authors or other users in the community.

Organization of this work is as follows: we first describe the theory
behind the various algorithmic components of Atropos while
acknowledging that theoretical discussion is available elsewhere,
e.g. in previously cited literature.  This is followed by a thorough
discussion of implementation which, though often overlooked, is of
immense practical utility.  We then report evaluation results on
public T1 neuroimaging datasets that have reference gold/silver
standard labelings.  Finally, we provide a discussion of our results
and our open-source contribution in the context of the remainder of
this paper and of previous and future work.

%  We evaluate this
%method on BrainWeb 20 data and show that by relying upon accurate,
%high-resolution diffeomorphic image registration, Atropos ~is capable of
%reliably segmenting the BW20 whole head data into separate tissue
%classes for muscle, skin, skull and bone marrow as well as the
%standard three tissues, cerebrospinal fluid and gray and white matter.
%Atropos 's three-tissue performance compares favorably with the standard
%EM-MRF algorithm, FAST.  We also use the OASIS dataset to show that
%Atropos ~increases segmentation repeatability relative to FAST.  The
%evaluation data, the data-derived templates, the code and application
%itself are publicly available.
%}
%\end{comment}

\section{Theoretical Foundations for Atropos Segmentation}

Atropos encodes a family of Bayesian segmentation techniques that may be configured in an application-specific manner.  The theory underlying Atropos dates back 20$+$ years and is representative of some of the most innovative work in the field.  Although we synthesize some of the theoretical work in this section, we recommend to the interested reader consultation with the literature cited previously for further insight.

Bayes' theorem provides a powerful mechanism for making inductive
inferences assuming the availability of quantities defining the
relevant conditional probabilities, specifically the likelihood and
prior probability terms.  Bayesian paradigms for brain image
segmentation employ a user-selected observation model defining the
likelihood term and one or more prior probability terms.  The product
of likelihood(s) and prior(s) is proportional to the posterior
probability.  The likelihood term has been previously defined both 
parametrically (e.g. a Gaussian model) and non-parametrically
(e.g. Parzen windowing of the sample histogram).  The prior term, as
given in the literature, has often been formed either as
MRF-based or template-based.  An image segmentation solution in this
context is an assignment of one label to each voxel
\footnote{
In the classic 3-tissue segmentation case, each voxel in the brain region is assigned a label of `cerebrospinal fluid (csf)', `gray matter (gm)', or `white matter (wm)'. 
}
such that the posterior probability is maximized.  The next sections
introduce notation and provide a formal description of three essential
components in Bayesian segmentation, viz.
\begin{itemize}
  \item the likelihood or observation model(s),
  \item the prior probability quantities derived from a generalized
    MRF and template-based prior terms, and 
%, and a novel distance prior used for label propagation, and 
  \item the optimization framework for maximizing the posterior probability.
\end{itemize}
These components are common across most EM segmentation approaches.

\subsection{Notation}
Assume a field, $\mathcal{F}$, whose values are known at discrete
locations, i.e. sites, within a regular voxel lattice that makes up an
image domain, $\mathcal{I}$.  Note that $\mathcal{F}$ can be a scalar
field in the case of unimodal data (e.g. T1 image only) or a vector
field in the case of multimodal data (e.g. T1, T2, and proton density
images).  A specific set of observed values, denoted by $\mathbf{y}$,
are indexed at $N$ discrete locations in $\mathcal{I}$ by $i \in \{1,
2, \ldots, N\}$.  This random field, $Y = \{y_1, y_2, \ldots, y_N \}$,
serves as a discrete representation of an observed image's intensities.  A labeling
of this image, also known as a hard segmentation, assigns to each site
in $\mathcal{I}$ one of $K$ labels from the finite set $\mathcal{L} =
\{l_1, l_2, \ldots, l_K\}$.  Also considered a random field, this
discrete labeling is $X = \{x_1, x_2, \ldots, x_N\}$ where each
$x_i \in \mathcal{L}$.   We use $\mathbf{x}$ to denote a specific set of labels in
$\mathcal{I}$ and a valid, though not necessarily optimal, solution to the segmentation problem.

%Furthermore, it is assumed that $\mathcal{F}$ is a random field whose true state, $\mathcal{S}$, is `hidden' beneath its observed values but fully represented by $\mathcal{L}$.%
%\footnote{
%For example, in the classic 3-tissue segmentation problem $\mathcal{L}$ is defined as the set $ \{\mathrm{csf}, \mathrm{gm}, \mathrm{wm}\}$ and each voxel in the brain region is assigned one and only one of these three labels.  
%}

\subsection{Segmentation Objective Function} Atropos optimizes a class
of user-selectable maximization objective functions each of which may
be represented in a generic Bayesian framework, as described by \cite{Sanjay-Gopal1998}.  This framework
requires {\em likelihood models} and {\em prior models} which enter
into Bayes' formula,
\begin{equation}\label{eq:bayes}
p(\mathbf{x}|\mathbf{y})=\underbrace{
p(\mathbf{y}|\mathbf{x})}_{\text{Likelihood(s)}} \underbrace{
p(\mathbf{x})}_{\text{Prior(s)}}\frac{1}{\mathbf{y}}, 
\end{equation} 
where the normalization term, $1/\mathbf{y}$, is a constant that does
not affect the optimization \citep{Sanjay-Gopal1998}.
Given choices for likelihood models and prior probabilities, the Bayesian
segmentation solution is the labeling $\hat{\mathbf{x}}$ which
maximizes the posterior probability, i.e.
\begin{align} \hat{\mathbf{x}} = \argmax_{\mathbf{x}}
\left\{p(\mathbf{y}|\mathbf{x})p(\mathbf{x})\right\}.
\end{align} Similar to its algorithmic predecessors, Atropos employs
the EM framework \citep{Dempster1977} to find maximum likelihood
solutions to this problem.  The following sections detail
Atropos choices for the likelihood(s), prior(s) and EM strategy.

\subsection{Likelihood or Observation Models}
To each of the $K$ labels corresponds a single probabilistic model describing the variation of $\mathcal{F}$ over $\mathcal{I}$.  We denote this set of $K$ likelihood models as $\Phi = \{p_1, p_2, \ldots, p_K\}$.  Using the standard notation, $\mathrm{Pr}(S=s) = p(s)$, $\mathrm{Pr}(S=s|T=t) = p(s|t)$, we can define these voxelwise probabilities, $\mathrm{Pr}_k( Y_i = y_i | X_i = l_k ) = p_k(y_i|l_k)$, in either parametric or non-parametric terms.   Given its simplicity and good performance, in the parametric case, $p_k$ is typically defined as a normal distribution, i.e.
\begin{align}\label{eq:param}
  p_k\left(y_i|l_k\right) &= G\left(\mu_k;\sigma_k\right) \nonumber \\
                    &= \frac{1}{\sqrt{2\pi \sigma_k^2}}\exp\left( \frac{ -(y_i - \mu_k)^2 }{2\sigma_k^2} \right)
\end{align}
where the parameters $\mu_k$ and $\sigma_k^2$ respectively represent
the mean and variance of the $k^{th}$ model.  When $y_i$ is a vector
quantity, we replace the Euclidean distance by by Mahalanobis distance
and define multivariate Gaussian parameters via a mean vector,
$\boldsymbol{\mu}_k$, and covariance matrix, ${\bf \Sigma_k}$.

A common technique for the non-parametric variant is to define $p_k$ using Parzen windowing of the sample observation histogram of $\mathbf{y}$, i.e.
\begin{align} \label{eq:nonparam}
  p_k\left(y_i|l_k\right) &= \frac{1}{N_B} \sum_{j=1}^{N_B} \frac{1}{\sqrt{2\pi \sigma_j^2}}\exp\left( \frac{ -(y_i - c_j)^2 }{2\sigma_j^2} \right)
\end{align}
where $N_B$ is the number of bins used to define the histogram of the sample observations (in Atropos the default is $N_B = 32$) and $c_j$ is the center of $j^{th}$ bin in the histogram.  $\sigma_j$ is the width of each of the $N_B$ Gaussian kernels.  For multi-modal data in which the number of components of $y_i$ is greater than one, a Parzen window function is constructed for each component.  The likelihood value is determined by the joint probability given by their product.

Atropos segmentation likelihood estimates are based on the classical finite mixture model (FMM).
FMM assumes independency between voxels to calculate the probability
associated with the entire set of observations, $\mathbf{y}$.  Spatial
interdependency between voxels is modeled by the prior probabilities
discussed in the next section.  Marginalizing over the set of possible labels, $\mathcal{L}$, leads to the following probabilistic formulation
\begin{align}\label{eq:likelihood}
  p(\mathbf{y}|\mathbf{x}) = \prod_{i=1}^N \left(      
                                                        \sum_{k=1}^K \gamma_k p_k(y_i|l_k)
                                                        \right)
\end{align}
where $\gamma_k$ is the mixing parameter \citep{Ashburner2005}.

\subsection{Prior Probability Models}
By modeling $\mathcal{F}$ via the set of observation models $\Phi$,
this so called finite-mixture model could be used to produce a
labeling or segmentation \citep[e.g.][]{Wells1996}.  However, as
pointed out by \cite{Zhang2001}, exclusive use of the intensity profile produces a less than optimal
solution because spatial contextual considerations are ignored.  This
has been remedied by the introduction of a host of prior probability
models including those characterized by use of MRF theory and
template-based information.  For example, in the works of
\cite{Leemput1999a} and \cite{Weisenfeld2009}, the original global
prior term given in \cite{Wells1996} is replaced by the product of the
template-based and the MRF-based prior terms.  In addition to their
descriptions below, we discuss a third possible
prior/objective combination in the form of a (sparse) prior labeling which fixes specific points
of the segmentation and uses EM to propagate this information
elsewhere in the image.  
% extension of template-based Euclidean and geodesic distance label propagation.

\subsubsection{Generalized MRF Prior}
One may incorporate spatial coherence into the segmentation by
favoring labeling configurations in which voxel neighborhoods tend
towards homogeneity. This intuition is formally described by MRF
theory in which spatial interactions in voxel neighborhoods can be
modeled \citep{Li2001}.

We assume the random field introduced earlier, $X$, is an MRF characterized by a neighborhood system, $\mathcal{N}_i$, on the lattice, $\mathcal{I}$, composed of the neighboring sites of $i$.  This neighborhood system is both noninclusive, i.e. $i \notin \mathcal{N}_i$, and reciprocating, i.e. $i \in \mathcal{N}_j \Leftrightarrow j \in \mathcal{N}_i$.   As an MRF, $X$ also satisfies the positivity and locality conditions, i.e., 
 \begin{align} \label{eq:mrf}
  p(\mathbf{x}) > 0, \,\, \forall\mathbf{x} 
 \end{align}
and where $\mathbf{x}$ is any particular labeling configuration on $X$ (in
other words, any labeling permutation on $X$ is {\em a priori}
possible).  The MRF locality condition is then, 
\begin{align}
  p\left(x_i | x_{\mathcal{I}-\{i\}}\right) = p\left(x_i | x_{\mathcal{N}_i}\right),
\end{align}
where $x_{\mathcal{I}-\{i\}}$ is the labeling of the entire image lattice except at site $i$ and  $x_{\mathcal{N}_i}$ is the labeling of $\mathcal{N}_i$.  This locality property enforces solely local considerations based on the neighborhood system in calculating the probability of the particular configuration, $\mathbf{x}$.  Following these two assumptions, the Hammersley--Clifford theorem provides the basis for treating the MRF distribution (cf. Eqn. (\ref{eq:mrf})) as a Gibbs distribution \citep{Besag1974,Geman1984}, i.e.
\begin{align} \label{eq:gibbs}
p(\mathbf{x}) = Z^{-1} \exp\left(-U(\mathbf{x})\right).
\end{align}
$Z$ is a normalization factor known as the {\em partition function}
and $U(\mathbf{x})$ is the {\em energy function} which can take
several forms \citep{Li2001}. In Atropos, as is the case with many
other segmentation algorithms of the same family, we choose
$U(\mathbf{x})$ such that it is only composed of a sum over pairwise
interactions between neighboring sites across the image,%
\footnote{
Using a more expansive definition of  $U(\mathbf{x})$, 
\begin{align}
U(\mathbf{x}) = \sum_{i = 1}^N \left( V_i(x_i) + \beta \sum_{j \in \mathcal{N}_i} V_{ij}( x_i, x_j ) \right), 
\end{align}
would permit casting the other prior terms inside the definition of $U(\mathbf{x})$ in the form of the external field $V_i(x_i)$ but, for clarity purposes, we consider them separately. 
}
i.e.
\begin{align}\label{eq:U}
U(\mathbf{x}) = \beta \sum_{i = 1}^N \sum_{j \in \mathcal{N}_i} V_{ij}( x_i, x_j )
\end{align}
where $V_{ij}$ is typically defined in terms of the Kronecker delta,
$\delta_{ij}$, based on the classical Ising potential (also known as a Potts model) \citep{Besag1974}
\begin{align}
V_{ij}(x_i, x_j) &= \delta_{ij} \nonumber \\
                          &= \left\{
                          \begin{array}{ll}
                            1 & \text{if } x_i = x_j \\
                            0 & \text{otherwise}
                          \end{array}
                         \right.   
\end{align}
and $\beta$ is a granularity term which weights the contribution of the MRF prior on the segmentation solution.
Since Atropos allows for non-uniform neighborhood systems and systems in which not just the immediate face-connected neighbors are considered, we use the modified function also used in \cite{Noe2001}, which weights the interaction term by the Euclidean distance, $d_{ij}$,  between interacting sites $i$ and $j$ such that 
\begin{align}
  V_{ij} = \frac{\delta_{ij}}{d_{ij}}
\end{align}
so that sites in the neighborhood closer to $i$ are weighted more heavily than sites further away.

\subsubsection{Template-Based Priors}
A number of researchers have used template-based strategies to both ensure spatial coherence and incorporate prior knowledge in segmentation.  A common technique is to select labeled subjects from a population from which a template is constructed \citep[e.g.][which is also available in the ANTs toolkit]{Avants2010}.  Each labeling can then be warped to the template where the synthesis of warped labeled regions produces a prior probability map or prior label map encoding the spatial distribution of labeled anatomy which can be harnessed in joint segmentation/registration or, specifically, Atropos/ANTs strategies involving unlabeled subjects.

We employ the strategy given in \cite{Ashburner2005} in which the
stationary mixing proportions, $\mathrm{Pr}(x_i = l_k) = \gamma_k$
(cf. Eqn. (\ref{eq:likelihood})), 
describing the prior probability that label $l_k$ corresponds to a particular voxel, regardless of intensity, are replaced by the following spatially varying mixing proportions,
\begin{align}
\mathrm{Pr}(x_i = l_k) = \frac{\gamma_k t_{ik}}{\sum_{j=1}^K\gamma_j t_{ij}}.
\end{align}
The $t_{ik}$ is the prior probability value at site $i$ which was
mapped, typically by image registration, to the local image
from a template data set.  The user may also
choose mixing proportions equal to
\begin{align}
\mathrm{Pr}(x_i = l_k) = \frac{t_{ik}}{\sum_{j=1}^K t_{ij}}.
\end{align}
via the command line interface to the posterior formulation.
{\textcolor{red}{FIXME --- need explanation of heuristic ... how do we
  estimate the $\gamma$ without the $t_{ij}$ present?}

\subsubsection{Supervised Semi-Interactive Segmentation}
Brain segmentation methods have relied on user interaction for many
years \citep{Lim1989,Julin1997,Freeborough1997a,Yushkevich2006}.
Atropos is capable of benefitting from user knowledge via an
initialization and optimization strategy that depends upon a spatially
varying prior label image passed as input.  Rapid, sparse
labeling---with visualization provided by ITK-SNAP
(www.itksnap.org)---enables an interaction and execution processing
loop that can be critical to solving segmentation problems with
challenging clinical data in which automated approaches fail.  This
part of Atropos design is inspired by the interactive graph cut methods
pioneered by \cite{Boykov2001} and which has spawned many follow-up
applications.  The Atropos prior label
image pre-specificies the segmentation results at a subset of the
spatial domain by fixing the priors and likelihood (and, thus, the
posterior) at a subset of $\mathcal{I}$ to be $1$ for the known label
and $0$ for each other label at the same site.  The user-input
therefore not only initializes the algorithm, but also gives boundary conditions
that influence the EM solution outside of the known sites.  While the
graph-based min-cut max-flow solution is globally optimal for two
labels, only locally optimal approachs are available for 3 or more
classes.  Thus, in most practical applications, the EM approach is a
reasonable and efficient alterative to Boykov's solution.
Furthermore, as with similar algorithms, one may automate the
initialization process.

\begin{comment}
{
\subsubsection{Distance Prior for Label Propagation}

In order to provide a dense segmentation even in regions where the template-based prior probability is 0, we use the template-based prior label images or probability maps described previously to formulate an optional distance prior probability.
Consider a subset of $\mathcal{I}$, which we denote $R_k$, defined by $\mathrm{Pr}(x_i = l_k) > \mathrm{Pr}(x_i = l_j) \,\,\forall j \in K, j \neq k$ and $\mathrm{Pr}(x_i = l_k) > 0$. 
We also denote the boundary of $R_k$ as $\partial R_k$ and the `peak' as $\rho_{R_k}$ where $i$ and $j$ are sites in $R_k$:
\begin{align}
  \rho_{R_k} = \argmax_{i \in R_k} \left\{ \min_{j \in \partial {R_k}} d\left(i,j\right)\right\}.
\end{align}
In other words, the peak is an interior site (or multiple interior sites) in $R_k$ characterized as having the maximum distance of all sites to any corresponding closest point on the boundary of $R_k$.  We denote this max-min distance as $\Delta_k$.  This allows us to define the distance prior probability for label propagation as follows:
\begin{align}\label{eq:prop}
  \mathrm{Pr}(x_i = l_k) = \left\{
                                               \begin{array}{ll}
                                                \alpha_k \exp \left( -\frac{d(i, \partial R_k)}{\sigma_k} \right) & \text{if } i \text{ is outside the region } R_k \\
                                                \alpha_k + d\left(i, \partial R_k\right)\left(\frac{1.0 - \alpha_k}{\Delta_k}\right) & \text{otherwise}
                                                \end{array}
                                                \right.
\end{align}
The probability calculations are illustrated in Fig.~\ref{fig:distancePrior} and were designed so that the most interior point, or set of points, of region $R_k$ would have a distance prior probability of 1.0 which would linearly decrease within the region to the boundary with a user-selected probability value of $\alpha_k$.  Outside $R_k$, the distance prior exponentially decays with a decay rate governed by another user-selected parameter $\sigma_k$.  This permits a dense segmentation throughout the segmentation region of $\mathcal{I}$ even where the template prior probabilities are zero.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=150mm]{distancePrior.pdf}
\end{tabular}
\caption{\baselineskip 12pt \small Graphical representation of the distance prior probability calculation associated with label $l_k$.  This extension of the popular template-based prior probability strategy permits a dense labeling with the user specified mask. }
\label{fig:distancePrior}
\end{center}
\end{figure}

A crucial matter, particularly for propagating labels in the cortical
regions, is the choice of distance function.  One of the advantages to
surface-based methods is that distances can be calculated along the
surface mesh, thus discriminating between proximal cortical surface
points and those points which are situated close together due to
cortical folding.  In Atropos, we provide two possibilities---(1) a
geodesic distance function using a fast marching construction
\citep{Osher1988} of the distance function with Atropos's masked
region of interest and (2) a Euclidean distance function using the algorithm described in \cite{Maurer2003}.  The differences in label propagation results are illustrated in Fig.~\ref{fig:label_propagation}.

Figure~\ref{fig:neonate} illustrates the benefit of geodesic distance
priors in segmenting T2 MRI of the 
neonatal brain.  Geodesic distance priors may be used to reduce ambiguity at
tissue interfaces, where partial voluming leads to mixed intensity.  
In neonatal brain segmentation from T2 MRI, the
interface of gray matter (low intensity) and cerebrospinal fluid
(bright intensity) often matches that of white matter (medium intensity). 
However, the white matter class should never appear near the surface
of the brain.  Thus, the distance from the brain surface may be used
to disambiguate voxels such as these.  
In this case, the distance of a voxel from a known
or expected tissue location serves to differentiate it from classes
that have similar intensity signature.  The geodesic distance is therefore an
alternative to explicit models of partial volume \citep{Ruan2000,Ballester2002,Leemput2003}.

\begin{figure}
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=50mm]{sparse_labels.png} &
\includegraphics[width=50mm]{prop_euclidean.png} &
\includegraphics[width=50mm]{prop_geodesic.png} \\
(a) & (b) & (c) \\
\end{tabular}
\caption{\baselineskip 12pt \small  Illustration of label propagation differences between the Euclidean and geodesic approaches which
are particularly acute in sinuous structures such as the cortex.  (a) Given a mask, represented in white, and a sparse labeling,
Atropos can be used to propagate the labels in a dense manner throughout the masked region.  (b) This propagation can occur using Euclidean distances in the image space via an Euclidean distance transform which can cause labels to jump meaningful anatomical boundaries.   (c) Given the cortical geometry, a more intuitive approach would be to propagate the labelings in a geodesic manner using a product of the fast marching paradigm which can be used in a manner that respects anatomical boundaries.  }
\label{fig:label_propagation}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{sparse_labels.png} 
\caption{\baselineskip 12pt \small  Neonatal brain segmentation.  }
\label{fig:neonate}
\end{center}
\end{figure}

}
\end{comment}

\subsection{Optimization}
Atropos uses expectation maximization to find a locally optimal
solution for the user-selected version of the Bayesian segmentation
problem (cf. Eqn. (\ref{eq:bayes})).
After initial estimation of the
likelihood model parameters, EM iterates between calculation of the
missing optimal labels $\hat{\mathbf{x}}$ and subsequent re-estimation of the model
parameters by maximizing the expectation of the complete data
log-likelihood (cf. Eqn. (\ref{eq:likelihood})).  
%\begin{align}
%\log( p(\mathbf{y}) ) = \sum_{i=1}^N \log \left( \sum_{k=1}^K \gamma_k  p(\mathbf{y}|l_k)  \right)
%\end{align}
The expectation maximization procedure is derived in various
publications including \cite{Zhang2001} which yields the optimal mean
and variance (or covariance), but sets the mixing parameter $\gamma_k$
as a constant.  The Atropos implementation estimates $\gamma_k$ at
each iteration as in \cite{Ashburner2005}.
\footnote{
Due to the lack of parameters in the non-parametric approach, it is not technically an EM algorithm (as described in \cite{Wells1996}).  However, the same iterative strategy is applicable and is quite robust in practice as evidenced by the number of researchers employing non-parametric strategies (see the Introduction).
}  
When spatial coherence constraints are included as an MRF prior in Atropos, the optimal segmentation solution becomes intractable.%
\footnote{
Consider $N$ sites each with a possible $K$ labels for a total of $N^K$ possible labeling configurations.  Due to the large $K$ associated with $K>>1$ problems, exact optimization is even more problematic than for the traditional 3-tissue scenario.
}
Although many optimization techniques exist \citep[see the
introduction in][for a concise summary of the myriad optimization
possibilities]{Marroquin2002}---each with their characteristic
advantages and disadvantages in terms of computational complexity and
accuracy---Atropos uses the well-known {\em Iterated Conditional
  Modes} (ICM)  \citep{Besag1986} which is greedy, computationally
efficient and provides good performance.  The EM algorithm employed in
Atropos may therefore be written as a series of steps:
\begin{description}
\item[Initialization:] In all cases, the user defines the number of
classes to segment.  The simplest initialization is by the classic
K-means algorithm with only the number of classes as a prior.
Otherwise, the user must provide prior information for each class in
the form of either a single $n$-ary prior label image or a series of
prior probability images, one for each class.
\item[E-Step: ICM Label Update:] Given the initialization and fixed
model parameters, Atropos will evaluate the posterior probability for
each possible label and assign, to each voxel, the label that
maximizes the posterior.  {\bf \textcolor{red}{ FIXME more specific?}}.
\item[M-Step: Parameter Update:] We describe the parametric update.  Given fixed labels, we update the
  parameters by the FMM-defined likelihood.  We compute, for each of
  $K$ labels, the mean,
$$ \mu^{new}_k \leftarrow \frac{ \sum_{i=1}^N  y_i \gamma_k p_k(y_i|l_k)}{ \sum_{i=1}^N \gamma_k p_k(y_i|l_k) },
$$
variance {\bf \textcolor{red}{ FIXME should this be the full
    covariance matrix?}},
$$ \sigma^{new}_k \leftarrow  \frac{ \sum_{i=1}^N  (y_i - \mu_k ) \gamma_k p_k(y_i|l_k)}{ \sum_{i=1}^N  \gamma_k p_k(y_i|l_k) },
$$
and mixture parameters, 
$$ \gamma^{new}_k \leftarrow  \textcolor{red}{ FIXME .}
$$
This type of update strategy is known as {\em soft} EM.  Hard EM, in
contrast, only uses sites containing label $l_k$ to update model $k$ parameters.
A similar strategy is used in the multivariate and non-parametric cases.
\end{description}
The EM algorithm will iterate toward a local maximum.  We track
convergence of the algorithm by summing up the maximum posterior
probability at each site over the segmentation domain.  
%\begin{align}
%P(l_k | y_i, t_i, m_i,d_i) &= \frac{P(y_i|l_k ,t_i,m_i,d_i)   P(l_k|t_i,m_i,d_i)   }{P(y_i|t_i,m_i,d_i)}
%\end{align}
%
%
%An image, $I$, maps a domain, $\Omega$ into the
%positive real numbers, such that $I \colon \Omega \rightarrow
%\mathbb{R}^+$.  The goal of segmentation, in general, is to define the
%spatial distribution of a finite set of labels over this domain.  We
%denote the segmentation itself as $\eta \colon \Omega \rightarrow L$
%where $L = \{ L_1 = 1 , \cdots , L_N=N \}$, a set of integer indexed
%segmentation labels.  Note that $\eta$ may be formed as $\eta =
%\sum_{i=1}^{i=N} L_i \eta_i $ where $\eta_i$ is the binary
%segmentation for label $L_i$.  A prior estimate for the label image
%$\eta_i$ is here denoted $\eta^s_i$ with a complete set of priors
%denoted $\eta^s$.  The boundary of the binary segmentation -- where a
%$0/1$ transition edge exists -- is denoted $\partial \eta^s_i$, for
%the prior, and $\partial \eta_i$ for the label image.
%
%\subsection{Apocrita Theory} A general maximum a posteriori criterion
%for segmentation seeks,
%\begin{equation} {\hat \eta} = \argmax _{\eta} \Pr( \eta | I )(\x) =
%\Pr( I | \eta )(\x) \Pr(\eta)(\x),
%\end{equation} 
%where $I$ is the input image, ${\eta}$ represents the
%label set configuration taken from the set $L$, $\Pr$ is the
%probability, $\x$ is the spatial index and the optimal solution is
%$\hat \eta$.  The input image $I$, here, is an unlabeled T1 MRI
%indexed by the value $\x \in \Omega$ where $\Omega$ is the image's
%spatial domain.  This probability is composed of the likelihood (first
%term) and the prior (second term).  Atropos ~ uses a spatially varying
%likelihood term and a two-component prior term that takes into account
%both spatial distribution and label smoothness, the latter via a
%standard MRF prior.
%
%The likelihood term for a single label value $L_i \in L$ is,
%\begin{eqnarray} \Pr( I | \eta_i )(\x)=\frac{1}{Z_i} \exp(- \| I(\x) -
%\mu_i(\x) \|^2 / \sigma_i^2 ),
%\end{eqnarray} where $Z_i$ is a normalizing constant, $\mu(\x)$ is a
%spatially varying estimate of the tissue mean and $\sigma$ is a
%standard deviation.  The prior term is given by,
%\begin{eqnarray} \Pr(\eta_i)(\x) = \frac{1}{Z^\prime_i} \exp( -f( \x -
%\y_{\partial \eta^s_i} ) / \sigma_{\eta^s_i}^2 ) p(\eta_i |
%\eta^\aleph_i), \\ \notag f( \x - \y_{\partial \eta^s_i} ) = (1 -
%\eta^s_i (\x) ) \| \x - \y_{\partial \eta^s_i} \|,
%\end{eqnarray} where $p(\eta_i | \eta^N_i)$ is the MRF smoothness
%probability based on the local neighborhood $\aleph$, the $Z$ is a
%normalizing constant and $\y_{\partial \eta^s_i}$ is the nearest point
%to $\x$ on the boundary of this labeling. Atropos ~requires a user or
%template-defined $\eta^s_i$ and the standard deviation
%$\sigma_{\eta^s_i}$ if a non-unity spatial prior component is desired
%for that label.  The free parameters, that must be estimated
%iteratively, are therefore $\mu_i, \sigma_i$ and the label set itself
%$\hat \eta$, which defines the (locally) optimal spatial distribution
%of $L$ through $\Omega$.  Figure~\ref{fig:spatp} shows the
%distribution of the spatial prior as a function of the distance from
%prior-defined object boundary.  Note that $f$ may easily be varied for
%other applications or that fixed probability images may also be
%substituted here.  This choice of $f$ is motivated by the fact that it
%allows compressed storage of the priors in a single image, $\eta^s$,
%while also maintaining the ability to manipulate -- for each
%$\eta^s_i$ -- the spatial influence of the prior via
%$\sigma_{\eta^s_i}$.  Practically, this is especially valuable when,
%$N$, the number of labels, is large.

\section{Implementation}

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=160mm]{AtroposFlowchart.pdf}
\end{tabular}
\caption{\baselineskip 12pt \small Flowchart illustrating Atropos usage typically beginning with bias correction using the N4 algorithm.  
Initialization provides an estimate before the iterative segmentation portion of the algorithm in which the 
likelihood models for each class are tabulated from the current estimate followed by a recalculation of the 
posterior probabilities associated with each class.  Label propagation ensues to ensure a dense 
classification throughout the entire mask.   The multiple options associated with the different algorithmic 
components are indicated by the colored rounded rectangles connected to their respective core Atropos 
processes via curved dashed lines.  }
\label{fig:flowchart}
\end{center}
\end{figure}
\subsection{The Atropos User Interface}
As with other classes that comprise ANTs, Atropos uses the Insight Toolkit as a developmental foundation.  
This allows us to take advantage of the mature portions of ITK (e.g. image IO) and ensures the 
integrity of the ancillary processes such as those facilitated by the underlying statistical framework.  Although Atropos is 
publicly distributed with the rest of the ANTs package, we plan to contribute the Atropos segmentation
framework to the Insight Toolkit where it can be vetted and further
improved by other interested researchers.

An overview of the core components of Atropos can be gleaned, in part, from the flowchart depicted in 
Fig.~\ref{fig:flowchart}.  To provide a more intuitive interface without the overhead costs of a graphical user interface, a set of unique command line parsing classes were developed which can also provide insight to the functionality of Atropos.  
The short version of the command line help menu is given in Listing \ref{listing:command} which is invoked by typing `{\ttfamily Atropos -h}' at the command prompt.  Both short and long option flags are available and each option has its own set of possible values and parameters introduced in a more formal way in both previous discussion and related papers cited in the introduction.  Here we  describe these options from the unique perspective of  implementation.

\singlespacing 
\begin{command}
\lstsetcpp
\begin{lstlisting}
COMMAND: 
     Atropos

OPTIONS: 
     -d, --image-dimensionality 2/3/4
     -a, --intensity-image [intensityImage,<adaptiveSmoothingWeight>]
     -b, --bspline [<numberOfLevels=6>,<initialMeshResolution=1x1x...>,<splineOrder=3>]
     -i, --initialization Random[numberOfClasses]
                          KMeans[numberOfClasses]
                          Otsu[numberOfClasses]
                          PriorProbabilityImages[numberOfClasses,
                            fileSeriesFormat(index=1 to numberOfClasses) or vectorImage,
                            priorWeighting,<priorProbabilityThreshold>]
                          PriorLabelImage[numberOfClasses,labelImage,priorWeighting]
     -x, --mask-image maskImageFilename
     -c, --convergence [<numberOfIterations=5>,<convergenceThreshold=0.001>]
     -k, --likelihood-model Gaussian
                            HistogramParzenWindows[<sigma=1.0>,<numberOfBins=32>]
     -m, --mrf [<smoothingFactor=0.3>,<radius=1x1x...>]
     -o, --output [classifiedImage,<posteriorProbabilityImageFileNameFormat>]
     -u, --minimize-memory-usage (0)/1
     -w, --winsorize-outliers BoxPlot[<lowerPercentile=0.25>,<upperPercentile=0.75>,
                               <whiskerLength=1.5>]
                              GrubbsRosner[<significanceLevel=0.05>,<winsorizingLevel=0.10>]
     -e, --use-euclidean-distance (0)/1
     -l, --label-propagation whichLabel[sigma=0.0,<boundaryProbability=1.0>]
     -h 
     --help 
\end{lstlisting} 
\end{command}

\subsection{Initializing the Atropos Objective}
The Atropos algorithm has a number of parameters defined within
Listing \ref{listing:n4} and will function on 2, 3 or 4 dimensional
data.  However, the majority of the time, users will be concerned with
a smaller set of input parameters.  Here, we list the recommended input and
an example definition for each parameter:
\begin{description}
\item[Input images to be segmented:] If more than
  one input image is passed, then a multivariate model will be
  instantiated.  E.g.  {\ttfamily -a Image.nii.gz} for one image and
  {\ttfamily -a Image1.nii.gz -a Image2.nii.gz} for multiple images.
\item[Input image mask:] This binary image
  will define the spatial segmentation domain.  Regions outside the
  mask receive the label $0$.  E.g.  {\ttfamily -x mask.nii.gz}. 
\item[Convergence criteria:] The algorithm terminates if it reaches max-iterations or
  produces a change less than the minimum change in the posterior.
  E.g. {\ttfamily -c [5,1.e-5]}. 
\item[MRF prior:] The key parameter to increase or decrease the
  spatial smoothness of the label map is $\beta$.  A useful range of
  $\beta$ values is $0$ to $0.5$ where we usually use $0.05, 0.1, 0.2$
  in brain segmentation.  E.g.  {\ttfamily -m [ 0.1, 1x1x1 ]} would
  define $\beta=0.1$ with a MRF radius of one voxel in each of three dimensions.
\item[Initialization:] The initialization options include (where the
  first parameter defines $K$, here 3 for each below),
\begin{itemize}
\item  {\ttfamily -i  kmeans[3]} standard kmeans initialization for
  three classes.
\item  {\ttfamily -i  PriorLabelImage[3, label\_image.nii.gz , w ]}
  where scalar $w \in [0,1]$ is unused \textcolor{red}{ FIXME .} 
\item  {\ttfamily -i  PriorProbabilityImages[3, label\_prob\%02d.nii.gz ,
    w ]} where $w=0$~(use the prior probability images only for
  initialization) or $w=0.5$ (use the prior probability images
  throughout the optimization).  If one chooses $0 < w < 0.5$ then one
  will increase (from zero) the weight on the priors.  These images,
  like the PriorLabelImage, should be defined with the same domain as
  the input images to be segmented.  
\end{itemize}
\item[Posterior formulation:]  The user may choose to estimate the
  mixture proportions (or not) by setting  {\ttfamily -p Socrates[1]}
  or {\ttfamily -p Socrates[0] }.  Fixed label boundary conditions may be employed by
  selecting the PriorLabelImage initialization and   {\ttfamily -p Plato[0] }.
\item[Output:] Atropos will output the hard segmentation and the
  probability image for each model.  E.g.  {\ttfamily -o
    [segmentation.nii.gz , seg\_prob\%02d.nii.gz] } will write
  out the hard segmentation in the first output parameter and a
  probability image for each class named, here, seg\_prob01.nii.gz ,
  seg\_prob02.nii.gz, etc. 
\end{description}
Higher dimensions than 4 are possible although we have not yet encountered such 
application-specific need.  Multiple images (assumed to be of the same
dimension, size, origin, etc.), will automatically enable multivariate
likelihoods.  In that case, the first image specified on the 
command line is used to initialize the Random, Otsu, or K-means labeling with the latter initialization further
refined by incorporating the additional intensity images, i.e. an initial univariate K-means clustering is determined 
from the first intensity image which, along with the other images, provides the starting multivariate cluster centers for a 
follow-up multivariate K-means labeling.    More details on each of
the key implementation options are given below.

\subsection{Likelihood Implementation}
As mentioned previously in the introduction, different groups have
opted for different likelihood models which have included either
parametric, specifically Gaussian, or nonparametric variations.
However, these approaches are similar in that they require a list
sample of intensity data from the input image(s) and a list of
weighting values for each observation of the list sample from which
the model is constructed.  In general, one may query model
probabilities by passing a given pixel's single intensity (for
univariate segmentation) or multiple intensities (for multivariate
segmentation) to the modeling function, regardless of whether the
function is parameter or non-parametric.  These similarities permitted
the creation of a generic plug-in architecture where classes
describing both parametric and nonparametric observational models are
all derived from an abstract list sample function class.  Three
likelihood classes have been developed, one parametric and two
nonparametric, and are available for usage although one of the
nonparametric classes is still in experimental development.  The
plug-in architecture even permits mixing likelihood models with
different classes during the same run for a hybrid
parametric/non-parametric approach although this possibility has yet
to be fully explored.

If the Gaussian likelihood model is chosen, the list sample of
intensity values and corresponding weights comprised of the posterior
probabilities are used to estimate the Gaussian model parameters, i.e.
the mean and variance.  For the non-parametric model, the list sample
and posteriors are used in a Parzen windowing scheme on a weighted
histogram to estimate the observational model \citep{Awate2006}.

\subsection{Prior Probability Models}
Consistent with our previous discussion, we offer both an MRF-based
prior probability for modeling spatial coherence and the possibility
of specifying a set of prior probability maps or a prior label map
with the latter extendable to creating a dense labeling.  To invoke
the MRF `{\ttfamily -m/--mrf}' option, one specifies the smoothing
factor (or the granularity parameter, $\beta$, given in
Eqn. (\ref{eq:U}), and the radius (in voxels) of the neighborhood
system using the vector notation `{\ttfamily 1x1x1}' for a
neighborhood radius of 1 in all 3 dimensions.  This radius is defined
such that voxels including but not limited to those that are face
connected will influence the MRF.

\begin{comment}{
The user also has an option to propagate initial labels, set by the
prior options, by a distance map tool.  
To propagate labels with the specification of prior probability images
or a prior label image, one needs to select the distance function (set
by the `{\ttfamily -e/--use-euclidean-distance}' boolean option)
which, by default, is determined by the geodesic distance.  The user
also needs to set the label propagation parameters for one or more of
the classes.   This  includes both the boundary probability and the exponential
 decay parameter, respectively $\alpha_k$ and $\sigma_k$ in Eqn. (\ref{eq:prop}).  
}\end{comment}

\subsection{Integration with Registration and N4 Bias Correction} 
Much of the image processing in our lab takes advantage of multiple
open source tools.  Most critically, these involve registration and
bias correction, in addition to segmentation. 
\subsubsection{Registration of Probability Maps} 
Registration approaches enable the ability to transform information
between spatial domains which may aid in both segmentation and bias
correction.  We rely heavily on template-building strategies
\citep{Avants2010,Avants2010a} which are also offered in ANTs.  Since
aligned prior probability images and prior label maps are often
associated with such templates, Atropos can be initialized with these
data with their influence regulated by a prior probability weighting
term.  Although prior label maps can be specified as a single
multi-label image, prior probability data are often represented as
multiple scalar images with a single image corresponding to a
particular label.  For relatively small classifications, such as the
standard 3-tissue segmentation (i.e. white matter, gray matter, and
cerebrospinal fluid), this does not typically present computational
complexities using modern hardware.  However, when considering dense
cortical parcellations where the number of labels can range upwards of
74 per hemisphere \citep{Destrieux2010}, the memory load can be
prohibitive if all label images are loaded into run-time memory
simultaneously.  A major part of minimizing memory usage in Atropos,
which corresponds to the boolean `{\ttfamily
-u/--minimize-memory-usage}' option, is the sparse representation of
each of the prior probability images.  Motivated by the observation
that these spatial prior probability maps tend to be highly localized
for large quantities of cortical labels, a threshold is specified on
the command line (default = 0.0) and only those probability values
which exceed that threshold are stored in the sparse representation.
During the course of optimization, the prior probability image for a
given label is reconstructed on the fly as needed.  For instance, the
NIREP (www.nirep.org) evaluation images are on the order of $300
\times 300 \times 256$ with 32 cortical labels.  Our novel memory
minimizing strategy typically shrinks run-time memory usage from a
peak of 10+ GB to approximately 1.5 GB and enable these datasets to be
used for training/prior-based cortical parcellation.

\singlespacing 
\begin{command}
\lstsetcppnfour
\begin{lstlisting}
COMMAND: 
     N4BiasFieldCorrection

OPTIONS: 
     -d, --image-dimensionality 2/3/4
     -i, --input-image inputImageFilename
     -x, --mask-image maskImageFilename
     -w, --weight-image weightImageFilename
     -s, --shrink-factor 1/2/3/4/...
     -c, --convergence [<numberOfIterations=50>,<convergenceThreshold=0.001>]
     -b, --bspline-fitting [splineDistance,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
                           [initialMeshResolution,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
     -t, --histogram-sharpening [<FWHM=0.15>,<wienerNoise=0.01>,<numberOfHistogramBins=200>]
     -o, --output [correctedImage,<biasField>]
     -h 
     --help 
\end{lstlisting} 
\end{command}
\subsubsection{Integration of N4 Bias Correction}
The typical segmentation processing pipeline begins with an intensity
normalization/bias correction step using the recently developed N4
algorithm \citep{Tustison2010} which is based on the popular
nonparametric nonuniformity intensity normalization (N3) method
introduced in \cite{Sled1998}.  It was previously mentioned that
several methods proposed in the literature have taken an integrative
view of the segmentation problem by incorporating an intrinsic bias
correction step into the actual algorithmic workflow.  The
advancements introduced with the N4 algorithm permit such an adaptive
integration with Atropos.  Recent demonstrations suggest improved
white matter segmentation produces better gain field estimates using
N3 \citep{Boyes2008}.  Thus, when performing 3-tissue segmentation, we
use the posterior probability map of white matter at the current
iteration as a weighted mask for input to N4.  This is done by setting
the `{\ttfamily --weight-image}' option on the N4 command line call
(see Listing \ref{listing:n4}) to the posterior probability image
corresponding to the white matter produced as output in the Atropos
call, i.e. `{\ttfamily Atropos --output}'.  The N4 algorithm is
currently being added as a standard part of the Insight ToolKit.  The
evaluation section will illustrate inclusion of Atropos, N4 and ANTs
in a brain processing pipeline.

\section{Evaluation}
Atropos~encodes a family of segmentation techniques that may be
instantiated for different applications but here we evaluate only two
of the many possibilites. First, we perform an evaluation on the
BrainWeb dataset using both the standard T1 image with multiple bias
and noise levels and also the BrainWeb20 data
\citep{Aubert-Broche2006a,Battaglini2008}.  In combination, this data
allows one to vary not only noise and bias but also the underlying
anatomy.  Second, we evaluate the use of Atropos in improving whole
brain parcellation and excercise its ability to efficiently solve {\em
many class} expectation maximization problem.  We choose this
evaluation problem in part to illustrate the flexibility of Atropos
and also the benefits of the novel, efficient implementation that
allows many-class problems to be solved with low memory usage ($<$2GB
for a 68 class model on 1mm$^3$ brain data).

\subsection{BrainWeb Evaluation}
\label{sec:bweb} The BrainWeb data is freely available at
\url{http://mouldy.bic.mni.mcgill.ca/brainweb/}.  We employ both the
individual subject data and the BrainWeb20 data in this evaluation.
\subsubsection{Single-Subject Evaluation} We use the single-subject
data with 3\% noise and bias 0, 20 and 40 \% RF inhomogeneity.  We
study the effect of the MRF prior term and initialization strategy on
the Dice overlap between ground truth and the segmentation result for
each tissue.  We test both K-means and prior label image
initialization with MRF $\beta \in \{ 0.00 , 0.05 , 0.10 , 0.15 , 0.20
, 0.25 , 0.30 \}$ at each bias field.  We also feed the white matter
probability map derived from K-means into N4 to guide the bias
correction.  Segmentation is then repeated, with the same parameters,
but with the N4 corrected image as input.  Thus, we are able to
compare the impact of bias on our N4 and Atropos algorithm and also
the validity of our prior label image initialization.  The prior label
image manual initialization was based on a crude ITK-SNAP-based labeling of a
small set of voxels for each tissue.  
\begin{figure}
\begin{center}
\begin{tabular}{c}
% \includegraphics[height=8in]{Figures/final_results_majority_vote_vs_atropos.pdf}
\end{tabular}
\caption{\baselineskip 12pt \small Brainweb single-subject results for
  each tissue.  The results show that the N4 bias correction
  algorithm, combined with Atropos, results in a minimal effect of
  bias, even at the 40\% level.  The optimal $\beta$ for the MRF term
  appears to be between $0.1$ and $0.2$.  The prior label image
  initialization is competitive with but inferior to K-means.}
\label{fig:hammers}
\end{center}
\end{figure}
\subsubsection{20 Subject Evaluation}

\subsection{The Hammers Dataset Evaluation}
\label{sec:hammers}
We evaluate the ability to improve multi-template
labeling results by converting the group labels to probability maps
and using them to drive many-class EM segmentation.  The procedure we
use here is similar to that described in \citep{Avants2010a}.  The
ground truth labels cover 68 classes and much of the brain.  Some
unlabeled regions remain which we assign to label 69 such that all
brain parenchyma contains a unique label.  The initialization of our
evaluation applies the script {\verb ants_multitemplate_labeling.sh }
(available in the ANTs toolkit) to the 19 Hammers evaluation
datasets located at \url{http://www.brain-development.org/} \citep{Hammers2003,Heckemann2006}.
These initial majority voting results are competitive with prior work
\citep{Heckemann2006,Heckemann2010} and serve as a baseline against
which we compare.

We first convert each of the 69 labels within the original evaluation
dataset to an individual image.  The remainder of the procedure,
summarized briefly, is the same for each of the 19 subjects.  We
select one subject as an unlabeled target.  The other 18 datasets are
then mapped (as in the script above) to that subject.  We then deform,
individually, the 69 $\times$ 18 label images to the unlabeled
subject.  The label probability map is gained by averaging the 18
deformed images associated with each label.  We repeat this for each
subject.  The following parameters are the most relevant to this discussion:
{\ttfamily  -i  PriorProbabilityImages[3, label\_prob\%02d.nii.gz , 0.5
  ] -m [0.1,1x1x1] -c [5,0] -p Socrates[1] }.  Results, in terms of
Dice overlap, are shown in figure~\ref{fig:hammers}.
\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[height=8in]{Figures/final_results_majority_vote_vs_atropos.pdf}
\end{tabular}
\caption{\baselineskip 12pt \small The figure compares the Dice
overlap results from Atropos versus the raw results from majority
voting for each of 68 neuroanatomical regions and, in addition, the
unlabeled portions of the brain from the Hammers evaluation dataset.
We evaluated Atropos via N-fold cross-validation and employed
PriorProbabilityImages for each class where probabilities are gained
by averaging mapped subject labels. The color coding highlights those
regions that have the highest (yellow) and lowest (pink) improvement.
The significance of the improvement, measured by pairwise T-test, is
also shown as is a trinary coding of that improvement as: $+$
significant improvement, $-$ performance reduction, $\sim$ no change.}
\label{fig:hammers}
\end{center}
\end{figure}

\section{Discussion}

{\textcolor{red}{ Elaborate after including the results.} }

A more general advantage which extends beyond the scope of the experimental 
evaluation section of this paper is the flexibility of the Atropos algorithm.   This includes 
not only $n$-tissue segmentation and dense volumetric cortical parcellation,
as reported in this work, but Atropos is also used in conjunction with our ANTs registration tools for robust
brain extraction which has reported good performance in comparison with other
popular, publicly available brain extraction tools \citep{Avants2010a}.  An additional successful
application area is that of ventilation-based segmentation of hyperpolarized helium-3 MRI
\citep{Tustison2010a}.

%\paragraph{Acknowledgments}
%{This work was supported in part by NIH ... }

\newpage

\bibliographystyle{neuroinformatics}
\bibliography{atropos}

\end{document}